---
title: "North Carolina"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
cache.lazy = FALSE
```

```{r, echo = FALSE}
library(tidyverse)
library(sf)
library(spData)
library(spDataLarge)
library(tmap)
library(leaflet)
library(RColorBrewer)
library(stringr)
library(stargazer)
```

```{r data setup, include = FALSE}
# Map Data 

county_gini <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/county_gini.csv")

nc_turnout <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/nc_turnout.csv")

nc_county <- st_read(dsn = "/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/nc_county20.shp")


county_gini_nc <- county_gini %>%
  filter(STATE == "North Carolina") %>%
  mutate(GEOID20 = str_c(STATEA, COUNTYA))

nc_county_gini <- left_join(nc_county, county_gini_nc, by = "GEOID20")
nc_county_gini <- left_join(nc_county_gini, nc_turnout, by = "NAMELSAD20")


# Voter File Data
nc_reg <- read_tsv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/ncvoter_Statewide.txt",n_max = 1000000)
nc_reg <- nc_reg[,c('voter_reg_num', 'voter_status_desc', 'voter_status_reason_desc', 'race_code', 'ethnic_code', 'party_cd', 'gender_code', 'birth_age', 'registr_dt', 'county_id')]

nc_vhis <- read_tsv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/ncvhis_Statewide.txt",n_max = 1000000)
nc_vhis <- nc_vhis[,c('county_id', 'county_desc', 'voter_reg_num', 'election_lbl')]

nc_voter <- left_join(nc_reg, nc_vhis, by = c('voter_reg_num' = 'voter_reg_num', 'county_id' = 'county_id'))

# Cleaning & Recoding

# Remove Deceased 
nc_voter <- filter(nc_voter, voter_status_desc != "REMOVED")
nc_voter <- filter(nc_voter, election_lbl == "11/03/2020" | is.na(election_lbl))
# Remove Old Votes


# Adding Gini
nc_gini <- county_gini_nc[,c("COUNTY", "AMEME001")]
nc_codes <- read.csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/nc_codes.csv")
nc_gini <- left_join(nc_gini, nc_codes, by = "COUNTY")
nc_voter <- left_join(nc_voter, nc_gini, by = "county_id")


# Vote 2020
nc_voter <-  mutate(nc_voter, vote2020 = case_when(nc_voter$election_lbl == "11/03/2020" ~ 1,
                           is.na(election_lbl) ~ 0))


# Registration Date (Month Before Election)
nc_voter$regdat <- as.Date(nc_voter$registr_dt, "%m/%d/%Y")
nc_voter <- mutate(nc_voter, mobilized = case_when(nc_voter$regdat > "2020-10-3" ~ 1,
                                                   regdat < "2020-10-3" ~ 0))


# Party Recode 

nc_voter <- mutate(nc_voter, democrat = case_when(nc_voter$party_cd == "DEM" ~ 1,
                                             race_code != "DEM" ~ 0))

nc_voter <- mutate(nc_voter, republican = case_when(nc_voter$party_cd == "REP" ~ 1,
                                             race_code != "REP" ~ 0))

nc_voter <- mutate(nc_voter, libertarian = case_when(nc_voter$party_cd == "LIB" ~ 1,
                                             race_code != "LIB" ~ 0))

nc_voter <- mutate(nc_voter, independent = case_when(nc_voter$party_cd == "UNA" ~ 1,
                                             race_code != "UNA" ~ 0))


# Gender 
nc_voter <- mutate(nc_voter, male = case_when(nc_voter$gender_code == "M" ~ 1,
                                             race_code != "M" ~ 0))
nc_voter <- mutate(nc_voter, female = case_when(nc_voter$gender_code == "F" ~ 1,
                                             race_code != "F" ~ 0))


# Race:
nc_voter <- mutate(nc_voter, asian = case_when(nc_voter$race_code == "A" ~ 1,
                                             race_code != "A" ~ 0))

nc_voter <- mutate(nc_voter, black = case_when(nc_voter$race_code == "B" ~ 1,
                                             race_code != "B" ~ 0))

nc_voter <- mutate(nc_voter, native = case_when(nc_voter$race_code == "I" ~ 1,
                                             race_code != "I" ~ 0))

nc_voter <- mutate(nc_voter, mixed = case_when(nc_voter$race_code == "M" ~ 1,
                                             race_code != "M" ~ 0))

nc_voter <- mutate(nc_voter, other = case_when(nc_voter$race_code == "O" ~ 1,
                                             race_code != "O" ~ 0))

nc_voter <- mutate(nc_voter, pacific = case_when(nc_voter$race_code == "P" ~ 1,
                                             race_code != "P" ~ 0))

nc_voter <- mutate(nc_voter, undesignated = case_when(nc_voter$race_code == "U" ~ 1,
                                             race_code != "U" ~ 0))

nc_voter <- mutate(nc_voter, white = case_when(nc_voter$race_code == "W" ~ 1,
                                             race_code != "W" ~ 0))
```

<p>&nbsp;</p>

# Introduction

This section zooms in to North Carolina as a case study in investigating our relationship of interest.  

Here, we are able to visualize the state in detail and leverage its publically available voter files to examine how county-level inequality influences propensity to vote while controlling for other factors.

# Data 

All data employed in this analysis was collected from the [North Carolina State Board of Elections](https://www.ncsbe.gov/) website in conjunction with ACS income inequality data.  

**NOTE:** North Carolina calculates turnout by the formula: $turnout = {votes\space cast\over registered \space voters}$, not traditional VEP/VAP standards

<p>&nbsp;</p>

# Looking at North Carolina

```{r}
library(tidyverse)
library(sf)
library(spData)
library(spDataLarge)
library(tmap)
library(leaflet)
library(RColorBrewer)
library(stringr)
library(stargazer)
```


##### Income Inequality by County
```{r incomemap, echo = FALSE}
tmap_mode("view")
tm_shape(nc_county_gini) +
  tm_polygons(col = "AMEME001", n = 8, palette ="BuGn", interactive = TRUE, id = "COUNTY", title = "Gini Coef.", legend.hist = TRUE, legend.is.portrait = FALSE, popup.vars = c("AMEME001")) +
  tm_layout(frame = FALSE, legend.outside = TRUE, legend.outside.position = "bottom") 
```
Interactive map displaying Gini Coefficients (2015-2019) per County.

<p>&nbsp;</p>
##### Voter Turnout by County
```{r turnoutmap, echo = FALSE}
tm_shape(nc_county_gini) +
  tm_polygons(col = "turnout", n = 8, palette ="Purples", interactive = TRUE, id = "COUNTY", title = "Voter Turnout (2020 General)", legend.hist = TRUE, legend.is.portrait = FALSE) +
  tm_layout(frame = FALSE, legend.outside = TRUE, legend.outside.position = "bottom") 
```
Interactive map displaying voter turnout (2020 General Election) per County.

<p>&nbsp;</p>
# Primary Analysis: Individual-Level Logit Model

The final analysis of the project is an individual-level analysis of North Carolina Voter Registration and Voter History files.  This logit model predicts individuals in North Carolina's propensity to vote in the 2020 General Election while controlling for key variables such as race, age, gender, party ID, and whether they registered to vote in the month preceding the election.

Before turning to the model, it is important to keep in mind a critical limitation of the model.  
- Due to technical limitations & difficulties, the sample included in this model is not randomly selected.  Rather, only the first million records in the voter registration and history files were selected to be analyzed.  
  - As a result, rather than randomly selecting, we have selected the files of voters from only 17 of the 100 counties in North Carolina, severely limiting variation in the main independent variable. Truthfully, this is a fatal flaw.

<p>&nbsp;</p>


##### The Model
```{r model, results = "asis"}

vote.logit <- glm(vote2020 ~ AMEME001 + democrat + libertarian + independent + female + asian + black + native + mixed + other + pacific + undesignated + birth_age + mobilized, family = binomial(link = "logit"), data = nc_voter, na.action = na.omit)

stargazer (vote.logit, type = 'html')
```

<p>&nbsp;</p>

##### Results
- Income inequality is positive & significant in predicting voting in the 2020 general election, holding all else at 0
- Libertarian & independent identification are negative & significant
- Black is the only racial category significant in predicting voting & is negatively related
- Age is positive & significant
- Registering to vote within a month of the election is negative & significant 
