---
title: "Analysis: State & County"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
library(tidyverse)
```

```{r data setup, include = FALSE}
# State Turnout
state_turnout <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/state_turnout_2020.csv") [-1,]
state_turnout <- state_turnout[,c('State', 'X6')]
names(state_turnout) <- c('state',
                          'vep')
state_turnout$vep <- as.numeric(sub("%","",state_turnout$vep))/100

# State Gini & Join

state_gini <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/state_gini.csv") 
state_gini <- state_gini[,c('STATE', 'AMEME001')]
names(state_gini) <- c('state', 'gini')

# State Join
states <- left_join(state_turnout, state_gini, by = 'state')
states <- na.omit(states)

# County Gini

county_gini <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/county_gini.csv")

# County Turnout

# Voters
countyvote <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/countypres_2000-2020.csv") %>%
  filter(year == "2020")

countyvote$county <- paste(countyvote$county_name, countyvote$state, sep = " County, ")

cvote <- countyvote %>%
	group_by(county) %>%
	summarise(vote = sum(candidatevotes))

cvote$county <- str_to_title(cvote$county)

# VAP Estimate
countyvap <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/CVAP_county.csv") %>%
    filter(lntitle == "Total")

countyvap <- rename(countyvap, GEOID = geoid)

countyvap <- countyvap[,c('GEOID', 'geoname', 'cvap_est')]
countyvap <- rename(countyvap, county = geoname)
countyvap$county <- str_to_title(countyvap$county)

county_turnout <- left_join(countyvap, cvote, by = "county")
county_turnout$turnout <- county_turnout$vote/county_turnout$cvap_est
county_turnout <- filter(county_turnout, turnout < 1)

# Adding County Ginis
iso_gini <- county_gini[,c('GEOID', 'AMEME001')]
county_turnout <- left_join (county_turnout, iso_gini, by = "GEOID")

```

# Intro

Previous literature suggests that the relationship between inequality and turnout is, to a degree, contingent on the level of analysis.  This section uses broad measures to examine this relationship at the state and county level, employing bivariate OLS models.  

The data employed here is the same as discussed in the previous "Visualizations" section.

<p>&nbsp;</p>

# State-Level Analysis

```{r}
library(tidyverse)
library(stargazer)
```


##### Visualizing State Inequality and Turnout (2020)
```{r, echo = FALSE}
ggplot(aes(x = gini, y = vep), data = states) +

    geom_smooth(method = "lm", fill = "gray", col = "red", se = TRUE) +
    labs(x = "Gini Coef.", y = "2020 VEP Turnout") +
    theme_minimal() +
    geom_text(aes(label = state), label.size = .1)
```
<p>&nbsp;</p>

##### The Model
```{r, results = "asis"}
state.lm <- lm(vep ~ gini, data = states)
stargazer(state.lm, type = "html", title = "Income Inequality and State-Level Turnout (VEP)", covariate.labels = "Gini Coef.")
```


<p>&nbsp;</p>
##### Interpretation

Consistent with the extant literature, we can see the relationship between inequality and voter turnout is negative at the state level (p value = 0.0576).

<p>&nbsp;</p>

# County-Level Analysis

<p>&nbsp;</p>

##### Visualizing County Inequality and Turnout (2020)

```{r, echo = FALSE}
ggplot(aes(x = AMEME001, y = turnout), data = county_turnout) +
    geom_point()+
    geom_smooth(method = "lm", fill = "gray", col = "red", se = TRUE) +
    labs(x = "Gini Coef.", y = "2020 Turnout") +
    theme_minimal()
```

<p>&nbsp;</p>

##### The Model
```{r, results = "asis"}
county.lm <- lm(turnout ~ AMEME001, data = county_turnout)
stargazer(county.lm, type = "html", title = "Income Inequality and County-Level Turnout (VAP Estimate)", covariate.labels = "Gini Coef.")
```
##### Interpretation

Surprisingly, the relationship holds even at the county level, which previous literature would not predict.  Moreover, while the coefficient is smaller than at the state level, this relationship is highly statistically significant.

<p>&nbsp;</p>

# Comparing

```{r, results = "asis", echo = FALSE}
stargazer(state.lm, county.lm, type = "html", covariate.labels = "Gini Coef.")
```

<p>&nbsp;</p>

##### Possible Explanations
1. The relationship between inequality and turnout at county-level is truly negative in the 2020 general election
2. This relationship could be the misleading artifact of a rough turnout measure for counties
3. This relationship could be the misleading artifact of comparing two different dependent variables (VEP vs. VAP)
