---
title: "Preliminary Visualizations"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)

```

```{r data setup, include = FALSE}
# State Turnout
state_turnout <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/state_turnout_2020.csv") [-1,]
state_turnout <- state_turnout[,c('State', 'X6')]
names(state_turnout) <- c('state',
                          'vep')
state_turnout$vep <- as.numeric(sub("%","",state_turnout$vep))/100

# State Gini & Join

state_gini <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/state_gini.csv") 
state_gini <- state_gini[,c('STATE', 'AMEME001')]
names(state_gini) <- c('state', 'gini')

# State Join
states <- left_join(state_turnout, state_gini, by = 'state')
states <- na.omit(states)

# County Gini

county_gini <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/county_gini.csv")

# County Turnout

# Voters
countyvote <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/countypres_2000-2020.csv") %>%
  filter(year == "2020")

countyvote$county <- paste(countyvote$county_name, countyvote$state, sep = " County, ")

cvote <- countyvote %>%
	group_by(county) %>%
	summarise(vote = sum(candidatevotes))

cvote$county <- str_to_title(cvote$county)

# VAP Estimate
countyvap <- read_csv("/Users/jakecaldwell/Jake/UF/Fall 2021/POS4931 - Election Data Science/Assignments/Project/CVAP_county.csv") %>%
    filter(lntitle == "Total")

countyvap <- rename(countyvap, GEOID = geoid)

countyvap <- countyvap[,c('GEOID', 'geoname', 'cvap_est')]
countyvap <- rename(countyvap, county = geoname)
countyvap$county <- str_to_title(countyvap$county)

county_turnout <- left_join(countyvap, cvote, by = "county")
county_turnout$turnout <- county_turnout$vote/county_turnout$cvap_est
county_turnout <- filter(county_turnout, turnout < 1)

# Adding County Ginis
iso_gini <- county_gini[,c('GEOID', 'AMEME001')]
county_turnout <- left_join (county_turnout, iso_gini, by = "GEOID")

```

<p>&nbsp;</p>

# Data 

The data used to generate the proceeding plots were collected from the following sources:  

1. State and county-level income inequality data (2015-2019) collected from American Community Survey.  
2. State-level voter turnout data collected from [ElectProject.org](http://www.electproject.org/) (should sound familiar).  
3. County-level turnout calculated by collecting raw county-level returns (MIT), divided by Census Bureau CVAP estimates.  This is a rather rough measure, with some important caveats.  Due to discrepancies in data entry, Alaska's counties were omitted, as well as a collection of smaller counties such as Dolores County, Colorado, which had voter turnout rates of over 100% for unknown reasons.  

<p>&nbsp;</p>
  
# Income Inequality by State and County

##### Measuring Inequality

Income inequality is generally measured via Gini coefficient, which ranges in value from 0 to 1.  A value of 0 suggests income is distributed entirely equally, while a value of 1 suggests that the entirety of the population's wealth is concentrated in one person.

```{r}
library(tidyverse)
```


```{r ineq, warning = FALSE, echo = FALSE}

state.ineq.hist <- ggplot(data = states, mapping = aes(x = gini)) +
  geom_histogram(binwidth = 0.01, color = "darkgreen", fill = "forestgreen") +
  geom_vline(xintercept = mean(states$gini), col = "darkblue", lwd = 1) +
  coord_cartesian(xlim = c(0.3, 0.65)) +
  labs(title = "State Inequality") +
  xlab("Gini Coefficient") +
  ylab("Count") +
  theme_minimal()

county.ineq.hist <- ggplot(data = county_turnout, mapping = aes(x = AMEME001)) +
  geom_histogram(binwidth = 0.01, color = "darkgreen", fill = "forestgreen") +
  geom_vline(xintercept = mean(county_turnout$AMEME001), col = "darkblue", lwd = 1) +
  coord_cartesian(xlim = c(0.3, 0.65)) +
  labs(title = "County Inequality") +
  xlab("Gini Coefficient") +
  ylab("Count") +
  theme_minimal()

grid.arrange(state.ineq.hist, county.ineq.hist, ncol = 2)
```

Here, we see that, unsurprisingly, there is much greater variation in the inequality among counties than there is among states.  

<p>&nbsp;</p>
<p>&nbsp;</p>

# Voter Turnout by State and County

```{r turn, warning = FALSE, echo = FALSE}
state.turn.hist <- ggplot(data = states, mapping = aes(x = vep)) +
  geom_histogram(binwidth = 0.015, color = "darkblue", fill = "navyblue") +
  geom_vline(xintercept = mean(states$vep), col = "forestgreen", lwd = 1) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(title = "State Turnout (VEP)") +
  xlab("VEP Turnout Rate") +
  ylab("Count") +
  theme_minimal()

county.turn.hist <- ggplot(data = county_turnout, mapping = aes(x = turnout)) +
  geom_histogram(binwidth = 0.015, color = "darkblue", fill = "navyblue") +
  geom_vline(xintercept = mean(county_turnout$turnout), col = "forestgreen", lwd = 1) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(title = "County Turnout (VAP Estimate)") +
  xlab("VAP Turnout Rate") +
  ylab("Count") +
  theme_minimal()

grid.arrange(state.turn.hist, county.turn.hist, ncol = 2)
```

Again, there is much greater variation among counties than states in voter turnout rates.  However, we should hesitate to draw any conclusions based on this graph.  These plots measure different phenomena - the former displaying the voting eligible population (VEP) turnout rate while the latter looks at voting age population (VAP) rate.  
